Localisation and Sensor Privacy Using the Extended Information Filter and Private Linear-Combination Aggregation\thanksref{footnoteinfo}

### 2020-10-13 main.tex
Started on making all the small changes listed in written notes Also fixing up English where its needed in first complete readthrough Making the changes suggested by Benjamin as they are reached - Currently up to LOC section, continuing from there tomorrow

\section{Introduction}
Localisation methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004} and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localisation methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large application due to their robustness and accuracy. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made these requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, a scenario of aircraft localisation in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Alternatively, autonomous vehicle state estimation may benefit from neighbouring vehicles' information while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localisation filter using this scheme such that sensor and navigator privacy is preserved.

@@

\subsection{Notation}
Throughout this work, we will use the following notation. Lowercase characters are scalars, $a$, vectors are underlined, $\vec{a}$, and random variables are bold, $\rv{a}$. Uppercase bold characters are matrices, $\mat{A}$, with $\mat{A}^{-1}$ denoting the inverse and $\mat{A}^\top$ the transpose. $|a|$ is the absolute value of $a$ and $\lVert\vec{a}\rVert$ the vector normal. $\E{\cdot}$ and $\Var{\cdot}$ are used for the expected value and variance, respectively. Encryption and decryption with key $k$ are denoted $\mathcal{E}_{k}(\cdot)$ and $\mathcal{D}_{k}(\cdot)$, respectively. $\lfloor\cdot\rceil$ rounds to the nearest integer, $a | b$ stands for ``$a$ divides $b$'', and $a\|b$ for a bitwise concatenation of $a$ and $b$. Sets will be written as $\{\dots\}$ while ordered sequences as $\langle\dots\rangle$.

@@

\section{Literature Overview}
In this section, we will discuss methods for providing security guarantees in signal processing tasks, and some applications to localisation and estimation problems.

@@

    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in aggregator obliviousness is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter mean calculation \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

Lastly, recent developments in encryption schemes have provided some novel methods also suitable for privacy-preserving computation in distributed environments.

@@

Relevant estimation methods include both model-based Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter and later derivatives \cite{kalmanNewApproachLinear1960,mutambaraDecentralizedEstimationControl1998}, and model-free methods where location estimates are made independently of each other \cite{alanwarPrOLocResilientLocalization2017,lazosSeRLocSecureRangeindependent2004}. The work in \cite{alanwarPrOLocResilientLocalization2017} uses partially homomorphic encryption to compute model-free location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided.

Model-based approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In \cite{aristovEncryptedMultisensorInformation2018}, partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements. Model-based encrypted control \cite{farokhiPrivacyDynamicalSystems2020} also introduces some privacy-preserving methods relevant to estimation. \cite{alexandruEncryptedCooperativeControl2019} proposes secure distributed control input aggregation using a novel weighted sum aggregation scheme, but requires repeated key distribution among all involved parties. In \cite{farokhiSecurePrivateControl2017}, a secure cloud-based control aggregator using partially homomorphic encryption is proposed, but is suitable only for untrusted service-providing cloud architectures.

While we have focused on the application of encryption schemes to localisation algorithms, homomorphic encryption schemes also require suitable number encoding when used with real numbers. As is the case with localisation sensors, real number measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. The Google BigNum library \cite{Encryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative for encoding is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings.

@@

The navigator state is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \rv{dx} & \rv{y} & \rv{dy}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $w_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$, while the measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
is defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}_k) &= \left\lVert
        \begin{bmatrix}
            \rv{x}_k & \rv{y}_k
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}_k-s_{x,i})^2 + (\rv{y}_k-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn the estimated state $\rvec{x}_k$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$ or their measurements $\rv{z}_k$ at any time $k$. We motivate these goals with an example. When considering aircraft navigation in the presence of privately-owned range-measuring towers, it is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown tower-owning parties. Similarly, tower locations may wish to be kept private from unidentified navigating aircraft. However, the additional safety to passengers, provided by accurate aircraft localisation, may be a goal of all those involved.

@@

\subsection{Problem Decomposition} \label{subsec:problem_decomposition}
The recursive nature of the problem definition above complicates defining and proving of security requirements for the localisation scenario. As with existing estimation methods [] we will reduce the communication requirements, of sensors and the navigator in our case, to an aggregation operation which will make security more easily definable. To support multiple aggregations at each time step $k$, as is required for the aggregation of multi-dimensional data, aggregation instances have been denoted by $t$. The bi-direction communication protocol at instance $t$ consists of of the computation of linear combinations of navigator weights by sensors, before their privacy-preserving aggregation by the navigator. That is, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are broadcast by the navigator, linear combinations $y^{(t)}_i=\sum^m_{j=1}x_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator. This has been summarised in Figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and combination step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle y_1^{(t)} \!=\! \sum^m_{j=1}x_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle y_n^{(t)} \!=\! \sum^m_{j=1}x_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (3.375,3.375) rectangle (5.625,2.625);
    \node at (4.5,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};
    \fill [lightgray] (0.875,3.375) rectangle (3.125,2.625);
    \node at (2,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

\begin{rem}
    From the aggregation definition above and the distributive property of real numbers, it is clear that an alternative method to the bi-directional communication requirement is computing $\sum^{m}_{j=1}\omega_j^{(t)}\sum^{n}_{i=1} x_{i,j}^{(t)}$ by the navigator, requiring only uni-directional communication by sending values $x_{i,j}^{(t)}, 1 \leq j \leq m$ from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in Section \ref{subsec:algorithm}.
\end{rem}

@@

\subsection{Participant Capabilities} \label{subsec:capabilities}
Before we can define concrete security requirements and a filtering protocol, we require some assumptions on the capabilities of the navigator and sensors in our problem scenario.
\begin{description}
    \item[Global navigator broadcast] Due to sensor location privacy and model non-linearity, bi-directional communication between navigator and sensors is required. We make the assumption that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.

    A problem caused by this requirement is apparent in the context of aircraft navigation. It may be the case that some sensors are not within the range of a navigator broadcast. We loosely propose that this scenario can be handled by pre-defined sensor subsets during an initial step by a trusted party. This has been captured in Figure \ref{fig:sensor_subsets}.
    \item[Consistent navigator broadcast] Further, we make the assumption that broadcast information from the navigator is received equal by all sensors. This means the navigator may not send different information to individual sensors during a single timestep.

@@

    % Gray range circles
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \draw[gray!50] (s1) circle (\srange);
        \draw[gray!50] (s2) circle (\srange);
        \draw[gray!50] (s3) circle (\srange);
        \draw[gray!50] (s4) circle (\srange);
        \draw[gray!50] (s5) circle (\srange);
    \end{scope}
    % Red subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[red!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Blue subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \fill[blue!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Magenta double overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[magenta!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Sensors
    \fill[black] (s1) circle (\sradius);
    \fill[black] (s2) circle (\sradius);
    \fill[black] (s3) circle (\sradius);
    \fill[black] (s4) circle (\sradius);
    \fill[black] (s5) circle (\sradius);
    % Bounding box
    \draw[gray] (brect_start) rectangle (brect_end);
    % Navigator
    \clip (brect_start) rectangle (brect_end);
    \pic[rotate=-40,xscale=0.22,yscale=0.3] at (nav_loc) {plane};
    \begin{scope}[shift={(nav_loc)},rotate=-40,xscale=0.22,yscale=0.3]
        % Trail
        \fill[bottom color=white, top color=gray] (-0.05, -1.5) rectangle (0.05, -50);
    \end{scope}
    % Sensor labels
    \node[label=1] at (s1) {};
    \node[label=2] at (s2) {};
    \node[label=3] at (s3) {};
    \node[label=4] at (s4) {};
    \node[label=5] at (s5) {};
\end{tikzpicture}
\caption{An example of sensor subsets; $\{1,2,3\}$ and $\{3,4,5\}$, and the areas where the navigator is in range of all sensors within a subset.}
\label{fig:sensor_subsets}
\end{figure}

@@

\subsection{Inherent Leakage}
Leakage of information will be discussed in more detail in Section \ref{subsec:leakage}, however, we stress that in any estimation problem where process and measurement models are known, knowledge of sequential state estimates naturally leads to the leakage of some measurement information. For this reason, we accept that average information about sensors may be leaked at each time step and instead show that individual sensor information remains private.

@@

\section{Cryptography Preliminaries} \label{sec:crypto_prelim}
When defining our aggregation security requirements and encryption scheme, we will reference some existing cryptographic security notions, the additively homomorphic Paillier encryption scheme, and the Joye-Libert private aggregation scheme.

@@

In addition to encryption and decryption, the following homomorphic functions are provided by the Paillier scheme. $\forall m_1,m_2 \in \mathbb{Z}_N$,
\begin{align}
    \mathcal{D}(\mathcal{E}(m_1)\mathcal{E}(m_2) \hspace{-7pt} \pmod{N^2}) &= m_1+m_2 \hspace{-7pt} \pmod{N} \\
    \mathcal{D}(\mathcal{E}(m_1)g^{m_2} \hspace{-7pt} \pmod{N^2}) &= m_1+m_2\hspace{-7pt}\pmod{N} \\
    \mathcal{D}(\mathcal{E}(m_1)^{m_2} \hspace{-7pt} \pmod{N^2}) &= m_1m_2 \hspace{-7pt} \pmod{N}\,. \label{eqn:paillier_hom_mult}
\end{align}

@@

\subsection{Joye-Libert Private Aggregation Scheme} \label{subsec:joye_libert_scheme}
The Joye-Libert private aggregation scheme \cite{joyeScalableSchemePrivacyPreserving2013} is a scheme defined on time-series data and meets the security notion of AO. Similarly to the Paillier scheme, it bases its security on the DCRA. A notable difference to a public-key encryption scheme is the need for a trusted party to perform an initial key generation and distribution step.

>>>>

\section{Introduction}
Localisation methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004} and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localisation methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large application due to their robustness and accuracy. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made the requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, an aircraft localisation scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighbouring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localisation filter using this scheme such that sensor and navigator privacy is preserved.

@@

\subsection{Notation}
Throughout this work, we will use the following notation. Lowercase characters are scalars, $a$, vectors are underlined, $\vec{a}$, and random variables are bold, $\rv{a}$. Uppercase bold characters are matrices, $\mat{A}$, with $\mat{A}^{-1}$ denoting the inverse and $\mat{A}^\top$ the transpose. $|a|$ is the absolute value of $a$ and $\lVert\vec{a}\rVert$ the vector normal. $\E{\cdot}$ and $\Var{\cdot}$ are used for the expected value and variance, respectively, and encryption and decryption with key $k$ are denoted $\mathcal{E}_{k}(\cdot)$ and $\mathcal{D}_{k}(\cdot)$. Additionally, key $k$ may be ommited when inferrable from context. $\lfloor\cdot\rceil$ rounds to the nearest integer, $a | b$ stands for ``$a$ divides $b$'', and $a\|b$ for a bitwise concatenation of $a$ and $b$. Sets will be written as $\{\dots\}$ while ordered sequences as $\langle\dots\rangle$.

@@

\section{Literature Overview}
In this section, we will discuss methods for providing security guarantees in signal processing tasks, and their applications to localisation and estimation problems.

@@

    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sum or mean calculation \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

Lastly, recent developments in encryption schemes have presented some novel methods also suitable for privacy-preserving computation in distributed environments.

@@

Relevant estimation methods include both Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter and later derivatives \cite{kalmanNewApproachLinear1960,mutambaraDecentralizedEstimationControl1998}, and time-independent methods where location estimates are made independently of each other \cite{alanwarPrOLocResilientLocalization2017,lazosSeRLocSecureRangeindependent2004}. The work in \cite{alanwarPrOLocResilientLocalization2017} uses partially homomorphic encryption to compute time-independent location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided.

Bayesian approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In \cite{aristovEncryptedMultisensorInformation2018}, partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements. Encrypted model predictive control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. \cite{alexandruEncryptedCooperativeControl2019} proposes secure distributed control input aggregation using a novel weighted sum aggregation scheme, but requires repeated key distribution among all involved parties. In \cite{farokhiSecurePrivateControl2017}, a secure cloud-based control aggregator using partially homomorphic encryption is proposed, but is suitable only for untrusted service-providing cloud architectures.

While we have focused on the application of encryption schemes to localisation algorithms, homomorphic encryption schemes also require suitable number encoding when used with real numbers. As is the case with localisation sensors, real number measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. Google's Encrypted BigQuery Client \cite{Encryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative for encoding is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings. We will discuss and justify our choice of encoding scheme when introducing the localisation filter.

@@

The navigator state is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \rv{dx} & \rv{y} & \rv{dy}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$, with measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
is defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}_k) &= \left\lVert
        \begin{bmatrix}
            \rv{x}_k & \rv{y}_k
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}_k-s_{x,i})^2 + (\rv{y}_k-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn the estimated state $\rvec{x}_k$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. However, the additional safety to passengers provided by accurate aircraft localisation, may be a goal of all those involved justifying their cooperation.

@@

\subsection{Problem Decomposition} \label{subsec:problem_decomposition}
The recursive nature of the problem definition above complicates defining and proving of security requirements for the localisation scenario. As with existing estimation methods \cite{benhamoudaNewFrameworkPrivacyPreserving2016,alexandruEncryptedCooperativeControl2019} we will reduce the communication requirements, of sensors and the navigator in our case, to an aggregation operation which will make security more easily definable.

To support multiple aggregations at each time step $k$, as is required for the aggregation of multi-dimensional data, aggregation instances have been denoted by $t$. The desired bi-direction communication protocol at instance $t$ consists of of the computation of linear combinations of navigator weights by sensors, before their privacy-preserving aggregation by the navigator. That is, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are broadcast by the navigator, linear combinations $y^{(t)}_i=\sum^m_{j=1}x_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator. This has been summarised in Figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and combination step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle y_1^{(t)} \!=\! \sum^m_{j=1}x_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle y_n^{(t)} \!=\! \sum^m_{j=1}x_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (3.375,3.375) rectangle (5.625,2.625);
    \node at (4.5,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};
    \fill [lightgray] (0.875,3.375) rectangle (3.125,2.625);
    \node at (2,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

\begin{rem}
    From the aggregation definition above and the distributive property of real numbers, it is clear that an alternative method to the bi-directional communication requirement would be computing $\sum^{m}_{j=1}\omega_j^{(t)}\sum^{n}_{i=1} x_{i,j}^{(t)}$ by the navigator, requiring only uni-directional communication by sending values $x_{i,j}^{(t)}, 1 \leq j \leq m$ from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in Section \ref{subsec:algorithm}.
\end{rem}

@@

\subsection{Participant Capabilities} \label{subsec:capabilities}
Before we can define concrete security requirements and a filtering protocol, we require some assumptions on the capabilities of the navigator and sensors in our problem scenario.
\begin{description}
    \item[Global navigator broadcast] From the decomposition defined above, we know that bi-directional communication between navigator and sensors is required. We make the assumption that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.

    A problem caused by this requirement is apparent in the context of aircraft navigation. It may be the case that some sensors are not within the range of a navigator broadcast. We loosely propose that this scenario can be handled with pre-defined sensor subsets during an initial step by a trusted party. An example has been shown in Figure \ref{fig:sensor_subsets}, which reduces to two separate localisation instances with global broadcasts.
    \item[Consistent navigator broadcast] Further, we make the assumption that broadcast information from the navigator is received equal by all sensors. This means the navigator may not send different information to individual sensors during a single timestep.

@@

    % Gray range circles
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \draw[gray!50] (s1) circle (\srange);
        \draw[gray!50] (s2) circle (\srange);
        \draw[gray!50] (s3) circle (\srange);
        \draw[gray!50] (s4) circle (\srange);
        \draw[gray!50] (s5) circle (\srange);
    \end{scope}
    % Red subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[red!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Blue subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \fill[blue!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Magenta double overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[magenta!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Sensors
    \fill[black] (s1) circle (\sradius);
    \fill[black] (s2) circle (\sradius);
    \fill[black] (s3) circle (\sradius);
    \fill[black] (s4) circle (\sradius);
    \fill[black] (s5) circle (\sradius);
    % Bounding box
    \draw[gray] (brect_start) rectangle (brect_end);
    % Navigator
    \clip (brect_start) rectangle (brect_end);
    \pic[rotate=-40,xscale=0.22,yscale=0.3] at (nav_loc) {plane};
    \begin{scope}[shift={(nav_loc)},rotate=-40,xscale=0.22,yscale=0.3]
        % Trail
        \fill[bottom color=white, top color=gray] (-0.05, -1.5) rectangle (0.05, -50);
    \end{scope}
    % Sensor labels
    \node[label=1] at (s1) {};
    \node[label=2] at (s2) {};
    \node[label=3] at (s3) {};
    \node[label=4] at (s4) {};
    \node[label=5] at (s5) {};
\end{tikzpicture}
\caption{An example of sensor subsets. Subsets $\{1,2,3\}$ and $\{3,4,5\}$ can be used independently for localisation when the navigator is within their ranges.}
\label{fig:sensor_subsets}
\end{figure}

@@

\subsection{Inherent Leakage}
Leakage of information will be discussed in more detail in Section \ref{subsec:leakage}, however, we stress that in any estimation problem where process and measurement models are known, knowledge of sequential state estimates as is the case at the navigator, or the knowledge of one or more measurements as is the case by one or more malicious sensors, naturally leads to the leakage of some information. For this reason, we accept that average sensor information may be leaked to the navigator and that less accurate estimates of location may be leaked to sensors. We will Instead show that individual sensor information remains private to sensors and that the most precise localisation estimate, requiring all measurements, remains private to the navigator.

@@

\section{Cryptography Preliminaries} \label{sec:crypto_prelim}
When defining our aggregation security requirements and encryption scheme, we will reference some existing cryptographic security notions, the additively homomorphic Paillier encryption scheme and the Joye-Libert private aggregation scheme.

@@

In addition to encryption and decryption, the following homomorphic functions are provided by the Paillier scheme. $\forall m_1,m_2 \in \mathbb{Z}_N$,
\begin{align}
    \mathcal{D}(\mathcal{E}(m_1)\mathcal{E}(m_2) \hspace{-7pt} \pmod{N^2}) &= m_1+m_2 \hspace{-7pt} \pmod{N} \\
    \mathcal{D}(\mathcal{E}(m_1)g^{m_2} \hspace{-7pt} \pmod{N^2}) &= m_1+m_2\hspace{-7pt}\pmod{N} \label{eqn:paillier_hom_plain_add}\\
    \mathcal{D}(\mathcal{E}(m_1)^{m_2} \hspace{-7pt} \pmod{N^2}) &= m_1m_2 \hspace{-7pt} \pmod{N}\,. \label{eqn:paillier_hom_mult}
\end{align}

@@

\subsection{Joye-Libert Private Aggregation Scheme} \label{subsec:joye_libert_scheme}
The Joye-Libert private aggregation scheme \cite{joyeScalableSchemePrivacyPreserving2013} is a scheme defined on time-series data, that we refer to as instances for consistency, and meets the security notion of AO. Similarly to the Paillier scheme, it bases its security on the DCRA. A notable difference to a public-key encryption scheme is the need for a trusted party to perform an initial key generation and distribution step.

Key generation is computed by choosing two equal length and sufficiently large primes $p$ and $q$, and computing $N=pq$. Additionally, hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$ is defined, and the public key is set to $(N, H)$. $n$ private keys are generated by choosing $sk_i,\,1\leq i\leq n$ uniformly from $\mathbb{Z}_{N^2}$ and distributing them to all $n$ users (whose values are to be aggregated), while the last key is set as
\begin{equation*}
    sk_0 = -\sum^{n}_{i=1}sk_i \pmod{N^2}\,,
\end{equation*}
and sent to the aggregator.
